---
title: "Haustoria TRAP-seq Analysis"
author: "Iliana Castillo-Machuca | Virginia Tech SPES"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
    theme: flatly
    highlight: tango
    df_print: paged
---

# 1. CLEAR ENVIRONMENT & SET PROJECT INFO
```{r CLEAR-ENVIRONMENT} 
#------------------------ CLEAN ENVIRONMENT ------------------------#
rm(list = ls(envir = .GlobalEnv, all.names = TRUE)) # Clear wrkspace objects
invisible(gc())  # Garbage collection to free memory

# Global options for consistent behavior
options(
  max.print              = 50000, # Cap console output
  scipen                 = 0,   # 999 to Avoid scientific notation, 0 if yes sci not
  stringsAsFactors       = FALSE, # Keep strings as character
  dplyr.summarise.inform = FALSE, # Suppress dplyr's messages
  digits                 = 4      # Number of digits to print
  )

set.seed(1234)
#------------------------ PROJECT INFO ------------------------# 
info <- list(
  Title             = "SPLIT GFP TRAP ANALYSIS",
  R_Version         = R.version$version.string,
  Working_Directory = getwd(),
  Timestamp         = format(Sys.time(), "%Y-%m-%d %H:%M:%S")
  )

separator <- paste0("\n", strrep("=", 80), "\n")
  cat(separator)
cat("PROJECT:", info$Title, "\n")
  cat(separator)
cat("R Version:", info$R_Version, "\n")
cat("Working Directory:", info$Working_Directory, "\n")
cat("Timestamp:", info$Timestamp, "\n")
  cat(separator)

#------------------------ SESSION INFO ------------------------#
si <- sessionInfo()

cat("Platform:  ", si$platform, "\n", sep = "")
cat("OS: ", si$running, "\n", sep = "")
  cat(separator)

#------------------------ REPRODUCIBILTY ------------------------#
#first time use to make an environment with the packages
#install.packages("renv")
#renv::init()

info$renv_active <- renv::project()
cat("renv Project:", info$renv_active, "\n")

info$Git_Commit <- system("git rev-parse --short HEAD", intern = TRUE)
cat("Git Commit:", info$Git_Commit, "\n")
```

# 2. INSTALL PACKAGES
```{r INSTALL-PACKAGES, include=TRUE, message=FALSE}
#------------------------ LOAD PACKAGES ------------------------#
cran_packages <- c(
  "tidyverse", # dplyr, ggplot2, tidyr, purrr, stringr, readr, forcats
  "patchwork", # Combine multiple ggplot objects
  "ggrepel", # Add non-overlapping text labels to ggplot2
  "pheatmap", # Clustered heat maps
  "RColorBrewer", # Color palettes
  "reshape2", # Data reshaping
  "dendextend", # Dendrogram
  "writexl", # Excel export
  "readxl",
  "tibble", # Tibbles
  "ggdendro",
  "knitr",
  "circlize",
  "rmarkdown"
  )
bioc_packages <- c(
  "DESeq2", # Differential expression analysis
  "AnnotationDbi", # Gene annotation framework
  "org.At.tair.db", # Arabidopsis thaliana annotations
  "clusterProfiler", # GO/KEGG enrichment analysis and GSEA
  "enrichplot", # Visualization of enrichment results
  "vidger", # RNA-seq viz helpers         
  "apeglm", # LFC shrinkage  
  "EnhancedVolcano",
  "ComplexHeatmap",
  "BiocManager",
  "GOSemSim"
)

#------------------------ INSTALL CRAN PACKAGES ------------------------#
new_cran <- cran_packages[!sapply(cran_packages, requireNamespace, quietly = TRUE)]
if (length(new_cran) > 0) install.packages(new_cran)

#------------------------ INSTALL BIOCONDUCTOR ------------------------#
if (!requireNamespace("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}

new_bioc <- bioc_packages[!sapply(bioc_packages, requireNamespace, quietly = TRUE)]
if (length(new_bioc) > 0) BiocManager::install(new_bioc, ask = FALSE)
 
#------------------------ LOAD PACKAGES ------------------------#
invisible(lapply(cran_packages, library, character.only = TRUE))
invisible(lapply(bioc_packages, library, character.only = TRUE))

#renv::status()
#renv::snapshot() 
```

# 3. SET DIRECTORIES
```{r SET-paths, include=FALSE}
#------------------------ 3.1 SET DIRECTORIES ------------------------# 
graphics_dir <- file.path("Results/Graphics")
tables_dir   <- file.path("Results/Tables") 
procdata_dir <- file.path("Results/Processed_data")

TRAPproj_dir <- file.path("~/Desktop/Lab/Haust_TRAP_proj/TRAP_Results")

dirs <- c(graphics_dir, tables_dir, procdata_dir, TRAPproj_dir) #this will be RDS obj for other projects

# Create directories if they don't exist
invisible(
  sapply(
    dirs,
    function(d) {
      if (!dir.exists(d)) {
        dir.create(d, recursive = TRUE, showWarnings = FALSE)
      }}
  ))
```

# 4. READ IN FILES
```{r READ-COUNTS, message=FALSE}
#------------------------ LOAD FEATURECOUNTS ------------------------# 
featcount <- read_tsv("../HvIM/Data_Repository/featcount.txt", skip = 1)
count_cols <- colnames(featcount)[7:ncol(featcount)]

read.count <- featcount %>%
  dplyr::select(all_of(count_cols)) %>%
  as.matrix()
rownames(read.count) <- featcount[[1]]

## Checks
stopifnot(!anyDuplicated(rownames(read.count)))
stopifnot(is.numeric(read.count[1, ]))
```

# 5. META DATA
```{r ALL-META-DATA, message=FALSE}
#------------------------ STANDARDIZE NAMES ------------------------# 
sample_names <- c("IT-107", "IT-108", "IT-109", "IT-110", 
                  "H-107", "H-108", "H-109", "H-110", "H-111",
                  "UT-106", "UT-108", "UT-114", 
                  "IT-206", "IT-207", "IT-208", "IT-209",
                  "IM-206", "IM-207", "IM-208", "IM-209", "IM-210",
                  "UT-204", "UT-208", "UT-213", "UT214",
                  "UM-204", "UM-208", "UM-213", "UM-214", "UM-215")
colnames(read.count) <- sample_names #save naming system

## Verify
cat("Counts loaded:", nrow(read.count), "genes x", ncol(read.count), "samples\n")

## Convert matrix to data frame with gene IDs
raw_counts_df <- as.data.frame(read.count)
raw_counts_df <- tibble::rownames_to_column(raw_counts_df, "GeneID")

# Export both CSV and Excel
#write_csv(raw_counts_df, file.path(tables_dir, "raw_counts_clean.csv"))
#writexl::write_xlsx(raw_counts_df, file.path(tables_dir, "raw_counts_clean.xlsx"))
```

# 6. EXTRACT H IM UM
```{r EXTRACT-count-data}
## Columns to DROP: IT(1:4,13:16), UT(10:12,22:25) -> KEEP 5-9,17-21,26-30
keep_cols <- c(5:9, 17:21, 26:30)
ImvHvUM <- read.count[, keep_cols]
colnames(ImvHvUM)  # Confirm order
cat("Counts loaded:", nrow(ImvHvUM), "genes x", ncol(ImvHvUM), "samples\n")

sample_info<- data.frame(
  row.names = c("H-107", "H-108", "H-109", "H-110", "H-111", 
                "IM-206", "IM-207", "IM-208", "IM-209", "IM-210", 
                "UM-204", "UM-208", "UM-213", "UM-214", "UM-215"),
  Groups = factor(c(rep("H",5), rep("IM",5), rep("UM",5)), levels=c("UM","IM","H")),
  Condition = factor(c(rep("Infected",10), rep("Uninfected",5)))
)

#checks
stopifnot(all(rownames(sample_info) == colnames(ImvHvUM)))

ImvHvUM_df <- tibble::rownames_to_column(as.data.frame(as.matrix(ImvHvUM)), "GeneID")

#write_csv(ImvHvUM_df, file.path(tables_dir, "counts_IM_H_UM_cleansubset.csv"))
#writexl::write_xlsx(ImvHvUM_df, file.path(tables_dir, "counts_IM_H_UM_cleansubset.xlsx"))

table(sample_info$Groups)
```

# 7. SET COLORS AND THRESHOLDS
```{r Colors}
#------------------------ SET COLOR AND THRESHOLDS ------------------------# 
colors <- c("H"="#CC79A7", "IM"="#7BC8A4", "UM"="#56B4E9", "UT"="#D55E00", "IT"="#F0E442")

pad <- 0.05
lfc_thresh <- 1.0
#------------------------ COLOR SWATCH ------------------------# 
df_colors <- data.frame(Condition = names(colors), Color = colors)

ggplot(df_colors, aes(x = Condition, y = 1, fill = Condition)) + geom_tile() + scale_fill_manual(values = colors) + theme_void() + theme( legend.position = "none", plot.margin = margin(2, 2, 2, 2)) + ggtitle("Color Swatch")
```

# 8. RUN DESeq
```{r DESeq-run, message = FALSE}
library(DESeq2)
#------------------------ SETUP DESEQ2 ------------------------#
dds <- DESeqDataSetFromMatrix(
  countData = ImvHvUM,
  colData = sample_info,
  design = ~ Groups)

## Prefilter low-count genes
keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep, ]
## UM reference
dds$Groups <- relevel(dds$Groups, ref = "UM")
cat("Counts loaded:", nrow(dds), "genes x", ncol(dds), "samples\n")

#------------------------ RUN DESEQ2 ------------------------#
dds <- DESeq(dds)

#------------------------ SAVE OBJECT ------------------------#
## Save DESeq object
#saveRDS(dds, file.path(TRAPproj_dir, "dds_IM_H_v_UM.rds"))

#------------------------ VIEW SIZE FACTORS ------------------------#
#sf <- round(sizeFactors(dds), 3)
#kable(data.frame(Sample=names(sf), SizeFactor=sf), caption="Size Factors")
```

# 9. DESEQ Results
```{r DESeq-results-summary}
#------------------------ RESULTS ------------------------#
res_IM_UM <- results(dds, name = "Groups_IM_vs_UM", alpha = 0.05)
res_H_UM  <- results(dds, name = "Groups_H_vs_UM", alpha = 0.05)

#------------------------ SAVE OBJECT ------------------------#
## Save DESeq object
#saveRDS(res_IM_UM, file.path(TRAPproj_dir, "raw_contrast_IM_v_UM.rds"))
#saveRDS(res_H_UM, file.path(TRAPproj_dir, "raw_contrast_H_v_UM.rds"))
```

# 10. VOLCANO PLOTS
```{r VOLCANO-PLOTS}
#------------------------ VOLCANO PLOTS ------------------------#
res_H_UM_df <- res_H_UM %>% as.data.frame() %>% rownames_to_column(var = "Gene") %>% filter(!is.na(padj)) %>% arrange(padj)
res_IM_UM_df <- res_IM_UM %>% as.data.frame() %>% rownames_to_column(var = "Gene") %>% filter(!is.na(padj)) %>% arrange(padj)
## Write Results
#writexl::write_xlsx(list(
#  "H_vs_UM" = res_H_UM_df,
#  "IM_vs_UM" = res_IM_UM_df),
#  file.path(tables_dir, "DESeq2_Results_pval.xlsx"))


# H vs UM volcano
EnhancedVolcano(
  res_H_UM_df,
  lab = res_H_UM_df$Gene,
  x = "log2FoldChange",
  y = "padj",
  pCutoff = 0.05,
  FCcutoff = 1,
  pointSize = 1.5,
  labSize = 3,
  title = "H vs UM",
  subtitle = paste("Total DE:", sum(res_H_UM_df$padj < 0.05, na.rm = TRUE)),
  caption = "Haustoria TRAP-seq"
)
#ggsave(file.path(graphics_dir, "Volcano_H_vs_UM.png"), width = 10, height = 8, dpi = 300)

# IM vs UM volcano  
EnhancedVolcano(
  res_IM_UM_df,
  lab = res_IM_UM_df$Gene,
  x = "log2FoldChange",
  y = "padj",
  pCutoff = 0.05,
  FCcutoff = 1,
  pointSize = 1.5,
  labSize = 3,
  title = "IM vs UM (infection response)",
  subtitle = paste("Total DE:", sum(res_IM_UM_df$padj < 0.05, na.rm = TRUE)),
  caption = "Infection Result TRAP-seq"
)
#ggsave(file.path(graphics_dir, "Volcano_IM_vs_UM.png"), width = 10, height = 8, dpi = 300)
```

# 11. VST NORMALIZATION
```{r VST-NORMALIZATION}
#------------------------ VST NORM ------------------------#
## Variance stabilizing transformation
vsd <- vst(dds, blind = FALSE)  # blind ignores design; subsample for speed

## Sample names, match to colData
colnames(vsd) <- rownames(sample_info)
stopifnot(all(colnames(vsd) == rownames(colData(dds))))  # Check

## Normalized counts matrix (size factors applied)
norm_counts <- counts(dds, normalized = TRUE)

# VST Summary Block
vsd_mat <- assay(vsd)  # Extract VST matrix
total_counts <- colSums(vsd_mat)

#print(summary(total_counts))

#------------------------ SAVE & EXPORT ------------------------#
#saveRDS(norm_counts, file.path(TRAPproj_dir, "ddsNORMcounts_H_IM_v_UM.rds"))
#saveRDS(vsd_mat, file.path(TRAPproj_dir, "ddsVSTcounts_H_IM_v_UM.rds"))

#EXPORT
#norm_counts_out <- norm_counts %>%
 # tibble::as_tibble(rownames = "id")
#VST_counts_out <- vsd_mat %>%
#  tibble::as_tibble(rownames = "id")

#write_csv(norm_counts_out,
 #         file.path(tables_dir, "normalized_counts_H_IM_UM.csv"))


#writexl::write_xlsx(
#  list(
#    Normalized = as.data.frame(norm_counts_out),  # Round for readability
#    VST = as.data.frame(VST_counts_out)), 
#  path = file.path(tables_dir, "H_IM_UM_normalized_VST.xlsx"))
#cat("✅ Exported Normalized + VST matrices to Tables/\n")
```

# 12. CLUSTERING & CORR
```{r clustering}
  ######################### Sample correlation heatmap (Spearman, raw counts >=1)  #########################
sim_mat <- ImvHvUM[rowSums(ImvHvUM) >= 1, ]
cor_samps <- cor(assay(vsd), method = "spearman")
melted_cor <- reshape2::melt(cor_samps)

## Correlation Plot
p_cor <- ggplot(melted_cor, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white", linewidth = 0.5) +
geom_text(aes(label = sprintf("%.2f", value)), color = "black", size = 3.2) +  scale_fill_gradient2(low = "white", high = "red", midpoint = 0.75, limits = c(0.5, 1), name = "Spearman ρ") +
  labs(title = "Sample Correlation (Raw Counts, Spearman)", x = "", y = "") +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(size = 9),
        panel.grid = element_blank(),
        legend.position = "bottom")
print(p_cor)
#ggsave(file.path(graphics_dir, "sample_correlation_spearman.png"), p_cor, width = 10, height = 8, dpi = 300)


  ######################### Hierarchical clustering (VST normalized, Euclidean distance)  #########################
sample_dist <- dist(t(assay(vsd)))  # t samples as rows
hc <- hclust(sample_dist, method = "ward.D2")
dend_data <- ggdendro::dendro_data(hc)


p_dend <- ggplot() +
  geom_segment(data = ggdendro::segment(dend_data), 
               aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_text(aes(x = x, y = y, label = label, colour = color), 
            hjust = -0.1, size = 3.5) +
  scale_colour_identity() +  # Use exact colors
  coord_flip() + 
  scale_y_reverse(expand = c(0.2, 0)) +
  ggdendro::theme_dendro() + 
  labs(title = "VST Clustering (Ward.D2)", 
       subtitle = "Labels colored by Groups (Haustoria/Infected Mesophyll/Uninfected Mesophyll)") +
  theme(plot.title = element_text(hjust = 0.5))
#print(p_dend)

#ggsave(file.path(graphics_dir, "sample_clustering_colored.png"), 
#       p_dend, width = 12, height = 8, dpi = 300)


  ######################### Sample distance matrix heatmap  #########################
dist_mat <- as.matrix(sample_dist)
colnames(dist_mat) <- rownames(sample_info) 
rownames(dist_mat) <- rownames(sample_info)

colors_dist <- colorRampPalette(rev(RColorBrewer::brewer.pal(9, "Blues")))(255)
p_dist <- pheatmap(dist_mat,
  clustering_distance_rows = sample_dist,
  clustering_distance_cols = sample_dist,
  color = colors_dist,
  main = "Sample Euclidean Distances (VST)",
  show_rownames = TRUE, show_colnames = TRUE,
  filename = file.path(graphics_dir, "sample_distance_heatmap.png"),
  width = 10, height = 9
)
```

# 13. PCA & PLOT
```{r PCA}
#------------------------ PCA PLOT ------------------------#
pca_data <- DESeq2::plotPCA(vsd, 
                           intgroup = c("Groups", "Condition"), 
                           returnData = TRUE)
percentVar <- round(100 * attr(pca_data, "percentVar"))

# Plot
p_pca <- ggplot(pca_data, aes(PC1, PC2, color = Groups, shape = Condition)) +
  geom_point(size = 4.5, stroke = 1.5, alpha = 0.85) +
  scale_color_manual(values = colors, name = "Sample Type") +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  geom_text_repel(aes(label = name),       
                      color = "black", 
                  box.padding = 0.35, 
                  max.overlaps = Inf,  # Show all labels
                  size = 3.5, 
                  segment.color = "red",
                  segment.size = 0.3,
                  bg.color = "white", 
                  bg.r = 0.1) +
  labs(title = "TRAP-seq Sample Clustering (Variance-Stabilized Counts)",
       subtitle = paste("PC1:", round(percentVar[1],1), "% variance |",
                       "PC2:", round(percentVar[2],1), "% variance"),
       caption = "Sample Type: UM (uninfected mesophyll), IM (infected mesophyll), H (haustoria)") +
  theme_classic(base_size = 14) +
  theme(legend.position = "bottom", 
        legend.title = element_text(face = "bold"),
        plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        plot.subtitle = element_text(hjust = 0.5, size = 14),
        panel.grid.minor = element_blank())
print(p_pca)

## EXPORT
#ggsave(file.path(graphics_dir, "TRAPseq_PCA_vst2.png"), p_pca, width = 8, height = 6, dpi = 300, bg = "white")
```

# 14. LFC SHRINKAGE + ARRANGEMENT + CLEAN
```{r DESeq-results-lfc, message=FALSE}
#------------------------ CONTRASTS ------------------------#
contrasts <- list(
  H_vs_UM  = list(coef = "Groups_H_vs_UM",  contrast = c("Groups", "H", "UM")),
  IM_vs_UM = list(coef = "Groups_IM_vs_UM", contrast = c("Groups", "IM", "UM"))
)

#------------------------ RUN RESULTS  ------------------------#
res_raw_list <- lapply(contrasts, function(x) {
  results(dds, contrast = x$contrast)
})

#------------------------ LFC Shrinkage  ------------------------#
# function to safely run lfcShrink with fallback
safe_lfcShrink <- function(dds, coef_name, contrast_vec) {
  tryCatch({
    lfcShrink(dds, coef = coef_name, type = "apeglm")
  }, error = function(e) {
    warning(paste0("apeglm failed for ", coef_name, ", using normal: ", e$message))
    lfcShrink(dds, contrast = contrast_vec, type = "normal")
  })
}
#------------------------ RUN Shrinkage  ------------------------#
lfc_list <- lapply(contrasts, function(x) {
  safe_lfcShrink(dds, x$coef, x$contrast)
})

#------------------------ SAVE OBJECT ------------------------#
# Save DESeq object
#saveRDS(lfc_list, file.path(TRAPproj_dir, "raw_shrunk_lfc_IM_H_v_UM.rds"))

#------------------------ CLEAN RESULTS  ------------------------#
#This section uses raw pvals and shrunk LFC
clean_res <- function(raw_res, shrink_res) {
  df <- as.data.frame(shrink_res)
  df$padj <- raw_res$padj
  df$pvalue <- raw_res$pvalue
  df$TAIR <- rownames(df)

  df %>%
    filter(!is.na(padj)) %>%
    arrange(padj)
}
#apply 
res_list <- mapply(clean_res, res_raw_list, lfc_list, SIMPLIFY = FALSE)

#------------------------ SAVE OBJECT ------------------------#
#saveRDS(res_list, file.path(TRAPproj_dir, "deseq_cleanshrunk_IM_H_v_UM.rds"))
```


```{r}
IM_H_VS_UM_load <- readRDS(file.path(TRAPproj_dir, "deseq_cleanshrunk_IM_H_v_UM.rds"))

dds_load <- readRDS(file.path(TRAPproj_dir, "dds_IM_H_v_UM.rds"))

vs_load <- readRDS(file.path(TRAPproj_dir, "ddsVSTcounts_H_IM_v_UM.rds"))
```

# 15. DEGs and THRESHOLD application
```{r SPLIT DEGS, message=FALSE}
#------------------------ ADD DEG CLASS ------------------------#
annotate_DEGs <- function(df, lfc_thresh, pad) {
  df %>%
   mutate(
      DEG = padj < pad & abs(log2FoldChange) >= lfc_thresh,
      direction = case_when(
        DEG & log2FoldChange >= lfc_thresh  ~ "UP",
        DEG & log2FoldChange <= -lfc_thresh ~ "DOWN",
        TRUE ~ "NS"
      )
   )}

#Apply
res_list <- lapply(IM_H_VS_UM_load, annotate_DEGs, lfc_thresh, pad) #res_list, not rds

#------------------------ ADD GENE ANNOTATION ------------------------#
add_annotations <- function(df) {
  df %>%
    mutate(
      symbol = mapIds(org.At.tair.db, keys = TAIR, column = "SYMBOL",
                      keytype = "TAIR", multiVals = "first"),
      entrez = mapIds(org.At.tair.db, keys = TAIR, column = "ENTREZID",
                      keytype = "TAIR", multiVals = "first")
    )}

#Apply
res_list <- lapply(res_list, add_annotations)

#------------------------ EXTRACT DEG SUBSETS ------------------------#
deg_list <- lapply(res_list, function(df) df %>% filter(DEG))
deg_split <- lapply(deg_list, function(df) split(df, df$direction))

#------------------------ SUMMARY TABLE ------------------------#
summary_tbl <- bind_rows(lapply(names(res_list), function(name) {
  df <- res_list[[name]]
  tibble(
    contrast = name,
    n_DEG  = sum(df$DEG),
    up     = sum(df$direction == "UP"),
    down   = sum(df$direction == "DOWN")
  )
}))
summary_tbl
#------------------------ FUNTION: SPLIT ENTREZ ------------------------#
entrez_list <- lapply(deg_split, function(contrast_list) {
  list(
    UP   = unique(contrast_list$UP$entrez[!is.na(contrast_list$UP$entrez)]),
    DOWN = unique(contrast_list$DOWN$entrez[!is.na(contrast_list$DOWN$entrez)])
  )
})

#------------------------ APPLY SPLIT ------------------------#
entrez_H_up   <- entrez_list$H_vs_UM$UP
entrez_H_down <- entrez_list$H_vs_UM$DOWN
entrez_IM_up   <- entrez_list$IM_vs_UM$UP
entrez_IM_down <- entrez_list$IM_vs_UM$DOWN
```

```{r}
deg_split

#------------------------ FUNTION: SPLIT ENTREZ ------------------------#
TAIR_list <- lapply(deg_split, function(contrast_list) {
  list(
    UP   = unique(contrast_list$UP$TAIR[!is.na(contrast_list$UP$TAIR)]),
    DOWN = unique(contrast_list$DOWN$TAIR[!is.na(contrast_list$DOWN$TAIR)])
  )
})

#------------------------ APPLY SPLIT ------------------------#
TAIR_H_up   <- TAIR_list$H_vs_UM$UP
TAIR_H_down <- TAIR_list$H_vs_UM$DOWN
TAIR_IM_up   <- TAIR_list$IM_vs_UM$UP
TAIR_IM_down <- TAIR_list$IM_vs_UM$DOWN
```


```{r MA PLOTS, message=FALSE}
#########################  MA PLOTS #########################
plot_DEGs_MA <- function(lfc_obj, df, title) {
  plotMA(lfc_obj, ylim = c(-5,5), main = title)
  
  sig <- df$DEG
  
  points(
    log2(lfc_obj$baseMean[sig] + 1),
    lfc_obj$log2FoldChange[sig],
    col = "red",
    pch = 16,
    cex = 0.6
  )
  
  invisible(NULL)
}

purrr::pwalk(
  list(lfc_list, res_list, names(res_list)),
  plot_DEGs_MA
)

#########################  EXPORT #########################
#export_list <- c(
#  res_list,
#  setNames(deg_list, paste0(names(deg_list), "_DEGs"))
#)

#writexl::write_xlsx(export_list, file.path(tables_dir, "DESeq2_results_shrink.xlsx"))
```


# 16. UP/DOWN splits for Hvs UM and IM vs UM
```{r Single Dotplot}
#------------------------ GO FUNCTION + SIMPLIFY ------------------------#
run_GO <- function(entrez_vec, ont, contrast_name, direction_name) {

  # Run GO
  go_raw <- enrichGO(
    gene          = entrez_vec,
    OrgDb         = org.At.tair.db,
    keyType       = "ENTREZID",
    ont           = ont,
    pAdjustMethod = "BH",
    pvalueCutoff  = 0.05,
    qvalueCutoff  = 0.05,
    readable      = TRUE
  )

  # simplify
  go_simplified <- tryCatch({
    simplify(go_raw, cutoff = 0.7, by = "p.adjust", select_fun = min)
  }, error = function(e) {
    message("simplify() failed for ", contrast_name, " ", direction_name,
            " (", ont, "). Returning unsimplified results.")
    go_raw
  })

  # metadata
  go_simplified@result$contrast  <- contrast_name
  go_simplified@result$direction <- direction_name
  go_simplified@result$ontology  <- ont

  return(go_simplified)
}

#------------------------ RUN GO ------------------------#
go_results <- list(
  H_up_BP    = run_GO(entrez_H_up,   "BP", "H_vs_UM",  "UP"),
  H_down_BP  = run_GO(entrez_H_down, "BP", "H_vs_UM",  "DOWN"),
  IM_up_BP   = run_GO(entrez_IM_up,   "BP", "IM_vs_UM", "UP"),
  IM_down_BP = run_GO(entrez_IM_down, "BP", "IM_vs_UM", "DOWN")
)

#MF
# go_results$H_up_MF    <- run_GO(entrez_H_up,   "MF", "H_vs_UM",  "UP")
# go_results$H_down_MF  <- run_GO(entrez_H_down, "MF", "H_vs_UM",  "DOWN")
# go_results$IM_up_MF   <- run_GO(entrez_IM_up,   "MF", "IM_vs_UM", "UP")
# go_results$IM_down_MF <- run_GO(entrez_IM_down, "MF", "IM_vs_UM", "DOWN")

# CC
# go_results$H_up_CC    <- run_GO(entrez_H_up,   "CC", "H_vs_UM",  "UP")
# go_results$H_down_CC  <- run_GO(entrez_H_down, "CC", "H_vs_UM",  "DOWN")
# go_results$IM_up_CC   <- run_GO(entrez_IM_up,   "CC", "IM_vs_UM", "UP")
# go_results$IM_down_CC <- run_GO(entrez_IM_down, "CC", "IM_vs_UM", "DOWN")


#------------------------ COMBINE (ALL) RESULTS ------------------------#
go_tbl <- bind_rows(lapply(go_results, function(x) x@result))
```


```{r}
#------------------------ COLLECT BP RESULTS ------------------------#
bp_list <- list(
  H_up    = go_results$H_up_BP@result,
  H_down  = go_results$H_down_BP@result,
  IM_up   = go_results$IM_up_BP@result,
  IM_down = go_results$IM_down_BP@result)

#------------------------ BASIC METRICS ------------------------#
bp_summary <- lapply(bp_list, function(df) {
  data.frame(
    contrast = unique(df$contrast),
    direction = unique(df$direction),
    n_terms = nrow(df),
    min_padj = min(df$p.adjust, na.rm = TRUE),
    median_padj = median(df$p.adjust, na.rm = TRUE),
    max_padj = max(df$p.adjust, na.rm = TRUE))
}) |> bind_rows()

#bp_summary

bp_tbl <- bind_rows(bp_list, .id = "group") %>%
  mutate(group = factor(group, levels = c("H_up","IM_up","H_down","IM_down")))

# Shared UP terms
shared_up <- intersect(
  bp_list$H_up$Description,
  bp_list$IM_up$Description)

# Shared DOWN terms
shared_down <- intersect(
  bp_list$H_down$Description,
  bp_list$IM_down$Description)

bp_up <- bp_tbl %>%
  filter(direction == "UP") %>%
  mutate(
    shared = ifelse(Description %in% shared_up, "shared", "unique"),
    shared = factor(shared, levels = c("unique", "shared")))

bp_down <- bp_tbl %>%
  filter(direction == "DOWN") %>%
  mutate(
    shared = ifelse(Description %in% shared_down, "shared", "unique"),
    shared = factor(shared, levels = c("unique", "shared")))

unified_scale <- scale_color_gradient(low = "hotpink", high = "#56B4E9")
```


```{r}
topN <- 20

bp_up_small <- bp_up %>% 
  group_by(group) %>%
  slice_min(order_by = p.adjust, n = topN) %>%
  ungroup()

bp_down_small <- bp_down %>%
  group_by(group) %>%
  slice_min(order_by = p.adjust, n = topN) %>%
  ungroup()

p_up <- ggplot(bp_up_small, aes( x = group, y = Description, size = Count, color = p.adjust, shape = shared)) +
  geom_point(alpha = 0.9) +
  unified_scale +
  scale_shape_manual(values = c("unique" = 17, "shared" = 16)) +
  ggtitle("GO BP — H & IM UP-Regulated Terms") +
  theme_bw(base_size = 12) +
  theme(
    plot.title  = element_text(face = "bold", size = 16, hjust = 0.5),
    axis.text.y = element_text(size = 12),
    strip.text  = element_text(size = 12, face = "bold")
  )
p_up

p_down <- ggplot(bp_down_small, aes(x = group, y = Description, size = Count, color = p.adjust, shape = shared)) +
  geom_point(alpha = 0.9) +
  unified_scale +
  scale_shape_manual(values = c("unique" = 17, "shared" = 16)) +
  ggtitle("GO BP — H & IM DOWN-Regulated Terms") +
  theme_bw(base_size = 12) +
  theme(
    plot.title  = element_text(face = "bold", size = 16, hjust = 0.5),
    axis.text.y = element_text(size = 12),
    strip.text  = element_text(size = 12, face = "bold")
  )
p_down

#ggsave(file.path(graphics_dir, "UP_Contrast_HvIM.png"), p_up, width = 10, height = 6, dpi = 300)
#ggsave(file.path(graphics_dir, "DOWN_Contrast_HvIM.png"), p_down, width = 10, height = 6, dpi = 300)
```


```{r}


TAIR_IM_up   <- TAIR_list$IM_vs_UM$UP
TAIR_IM_down
# ---- KEGG: H vs UM ----
ekegg_HUM_up <- enrichKEGG(
  gene          = TAIR_H_up,
  organism      = "ath",
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05
)

ekegg_HUM_down <- enrichKEGG(
  gene          = TAIR_H_down,
  organism      = "ath",
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05
)

# ---- KEGG: IM vs UM ----
ekegg_IMum_up <- enrichKEGG(
  gene          = TAIR_IM_up,
  organism      = "ath",
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05
)

ekegg_IMum_down <- enrichKEGG(
  gene          = TAIR_IM_down,
  organism      = "ath",
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05
)

# ---- Load required packages ----
library(clusterProfiler)
library(enrichplot)

# ---- Combine all KEGG enrichment results ----
combined_KEGG_all <- merge_result(list(
  HUM_up  = ekegg_HUM_up,
  IMUM_up = ekegg_IMum_up,
  HUM_down   = ekegg_HUM_down,
  IMUM_down   = ekegg_IMum_down
))

# ---- Visualize combined KEGG enrichment ----
dotplot(combined_KEGG_all,
        showCategory = 15,
        title = "Combined KEGG Enrichment: H vs UM and IM vs UM") +
  theme(axis.text.y = element_text(size = 10))
```



# 18. FOURWAY PLOT
```{r VIDGER}
#------------------------ FOUR-WAY PLOT ------------------------#
FW_df <- vsFourWay( x = "H", y = "IM", control = "UM", 
                    data = dds_load, 
                    d.factor = "Groups", 
                    type = "deseq", 
                    padj = pad, 
                    lfc = lfc_thresh, 
                    title = TRUE, 
                    grid = TRUE, 
                    legend = TRUE, 
                    data.return = TRUE, 
                    xaxis.text.size = 14, 
                    yaxis.text.size = 14, 
                    legend.text.size = 10) 
print(FW_df$plot) # View plot
#------------------------ SAVE ------------------------#
#ggsave(file.path(graphics_dir, "FourWay_plot.png"), FW_df$plot, width = 8, height = 6, dpi = 300)

#vignette("vidger")

```

```{r}
genes_of_interest <- c("AT3G09770", "AT4G31730", "AT2G36270", "AT2G40220") # order: LOG2, GDU1, ABI5, ABI4


# Get log2FC directly
res_H  <- results(dds, contrast = c("Groups", "H", "UM"))
res_IM <- results(dds, contrast = c("Groups", "IM", "UM"))

df <- data.frame(
  gene = rownames(res_H),
  x = res_H$log2FoldChange,
  y = res_IM$log2FoldChange
)

subset_df <- df[df$gene %in% genes_of_interest, ]

# Plot with highlighted genes
p <- FW_df$plot +
  geom_point(data = subset_df, aes(x = x, y = y),
             color = "hotpink", size = 3) +
  geom_text_repel(data = subset_df, aes(x = x, y = y, label = gene),
                  color = "hotpink")

print(p)

# Save
#ggsave(file.path(graphics_dir, "FourWay_H_vs_IM_vs_UM_LOG2_GDU1.png"), p, width = 10, height = 8, dpi = 300)
```


```{r}
fw_extract <- FW_df$data

H_specific <- fw_extract %>%
  filter(color == "green",
         padj_x < pad,
         abs(logFC_x) > lfc_thresh,
         padj_y >= pad)

H_specific_up <- H_specific %>% filter(logFC_x > 0)
H_specific_down <- H_specific %>% filter(logFC_x < 0)


IM_specific <- fw_extract %>%
  filter(color == "red",
    padj_y < pad,
         abs(logFC_y) > lfc_thresh,
         padj_x >= pad)

IM_specific_up <- IM_specific %>% filter(logFC_y > 0)
IM_specific_down <- IM_specific %>% filter(logFC_y < 0)


fourway_workbook <- list(
  H_up_green   = H_specific_up,
  H_down_green = H_specific_down,
  IM_up_red  = IM_specific_up,
  IM_down_red= IM_specific_down
)

#writexl::write_xlsx(fourway_workbook, file.path(tables_dir, "Green_red_extract.xlsx"))
```



```{r GO-GREEN-RED}
#------------------------ GO: GREEN - HAUSTORIA ------------------------# 
ego_H_up <- enrichGO(
  gene = H_specific_up$id,
  OrgDb = org.At.tair.db,
  keyType = "TAIR",
  ont = "BP",
  pAdjustMethod = "BH",
  pvalueCutoff = pad,
  qvalueCutoff = pad,
  minGSSize = 3)

ego_H_down <- enrichGO(
  gene = H_specific_down$id,
  OrgDb = org.At.tair.db,
  keyType = "TAIR",
  ont = "BP",
  pAdjustMethod = "BH",
  pvalueCutoff = pad,
  qvalueCutoff = pad,
  minGSSize = 3)

#------------------------ GO: RED - INFECTED MESOPHYLL ------------------------#  
ego_IM_up <- enrichGO(
  gene = IM_specific_up$id,
  OrgDb = org.At.tair.db,
  keyType = "TAIR",
  ont = "BP",
  pAdjustMethod = "BH",
  pvalueCutoff = pad,
  qvalueCutoff = pad,
  minGSSize = 3)

ego_IM_down <- enrichGO(
  gene = IM_specific_down$id,
  OrgDb = org.At.tair.db,
  keyType = "TAIR",
  ont = "BP",
  pAdjustMethod = "BH",
  pvalueCutoff = pad,
  qvalueCutoff = pad,
  minGSSize = 3)

#------------------------ SIMPLIFY ------------------------#
ego_H_up<- simplify(ego_H_up)
ego_H_down<- simplify(ego_H_down)
ego_IM_up<- simplify(ego_IM_up)
ego_IM_down<- simplify(ego_IM_down)

df_H_up  <- ego_H_up@result
df_IM_up <- ego_IM_up@result

df_H_down  <- ego_H_down@result
df_IM_down <- ego_IM_down@result

unified_scale <- scale_color_gradient(low = "#56B4E9", high = "hotpink")
shape_map <- scale_shape_manual(values = c("unique" = 17, "shared" = 16))
```


```{r}

# ---- KEGG: H vs UM ----
ekegg_HUM_up <- enrichKEGG(
  gene          = H_specific_up$id,
  organism      = "ath",
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05
)

ekegg_HUM_down <- enrichKEGG(
  gene          = H_specific_down$id,
  organism      = "ath",
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05
)

# ---- KEGG: IM vs UM ----
ekegg_IMum_up <- enrichKEGG(
  gene          = IM_specific_up$id,
  organism      = "ath",
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05
)

ekegg_IMum_down <- enrichKEGG(
  gene          = IM_specific_down$id,
  organism      = "ath",
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05
)

# ---- Load required packages ----
library(clusterProfiler)
library(enrichplot)

# ---- Combine all KEGG enrichment results ----
combined_KEGG_all <- merge_result(list(
  HUM_up  = ekegg_HUM_up,
  IMUM_up = ekegg_IMum_up,
  HUM_down   = ekegg_HUM_down,
  IMUM_down   = ekegg_IMum_down
))

# ---- Visualize combined KEGG enrichment ----
dotplot(combined_KEGG_all,
        showCategory = 15,
        title = "Combined KEGG Enrichment: H vs UM and IM vs UM") +
  theme(axis.text.y = element_text(size = 10))
```


```{r GO-UP-PREP}
topN <- 15

df_H_up_small <- df_H_up %>%
  arrange(p.adjust) %>%
  slice_head(n = topN) %>%
  mutate(group = "H_UP")

df_IM_up_small <- df_IM_up %>%
  arrange(p.adjust) %>%
  slice_head(n = topN) %>%
  mutate(group = "IM_UP")


shared_up_terms <- intersect(df_H_up_small$Description,
                             df_IM_up_small$Description)


df_up_combined_small <- bind_rows(df_H_up_small, df_IM_up_small) %>%
  mutate(
    shared = ifelse(Description %in% shared_up_terms, "shared", "unique"),
    shared = factor(shared, levels = c("unique", "shared"))
  )
```


```{r GO DOWN - PREP}
topN <- 15

df_H_down_small <- df_H_down %>%
  arrange(p.adjust) %>%
  slice_head(n = topN) %>%
  mutate(group = "H_DOWN")

df_IM_down_small <- df_IM_down %>%
  arrange(p.adjust) %>%
  slice_head(n = topN) %>%
  mutate(group = "IM_DOWN")


shared_down_terms <- intersect(df_H_down_small$Description,
                             df_IM_down_small$Description)


df_down_combined_small <- bind_rows(df_H_down_small, df_IM_down_small) %>%
  mutate(
    shared = ifelse(Description %in% shared_down_terms, "shared", "unique"),
    shared = factor(shared, levels = c("unique", "shared"))
  )
```



```{r}
p_up_combined <- ggplot(df_up_combined_small, aes(
  x = group,
  y = Description,
  size = Count,
  color = p.adjust,
  shape = shared
)) +
  geom_point(alpha = 0.9) +
  unified_scale +
  shape_map +
  ggtitle("GO BP — Combined H_UP & IM_UP (Shared vs Unique)") +
  theme_bw(base_size = 12) +
  theme(
    plot.title  = element_text(face = "bold", size = 16, hjust = 0.5),
    axis.text.y = element_text(size = 12)
  )
p_up_combined

p_down_combined <- ggplot(df_down_combined_small, aes(
  x = group,
  y = Description,
  size = Count,
  color = p.adjust,
  shape = shared
)) +
  geom_point(alpha = 0.9) +
  unified_scale +
  shape_map +
  ggtitle("GO BP — Combined H_DOWN & IM_DOWN (Shared vs Unique)") +
  theme_bw(base_size = 12) +
  theme(
    plot.title  = element_text(face = "bold", size = 16, hjust = 0.5),
    axis.text.y = element_text(size = 12),
    strip.text  = element_text(size = 12, face = "bold")
  )
p_down_combined
```



# 20. TRANSPORTERS, EXTRACT & SAVE
```{r}
#------------------------ LOAD TRANSPORTERS ------------------------#
transporter_genes <- read_excel("../HvIM/Data_Repository/6_transporter_df.xlsx") %>%
  rename(id = TAIR) %>%
  mutate(id = trimws(as.character(id)))  # make sure it's character, no spaces
transporter_genes$id <- gsub("\\..*", "", transporter_genes$id)

#CLEAN IDS in tables
H_specific_up <- H_specific_up %>% mutate(id = trimws(as.character(id)))
H_specific_down <- H_specific_down %>% mutate(id = trimws(as.character(id)))
IM_specific_up <- IM_specific_up %>% mutate(id = trimws(as.character(id)))
IM_specific_down <- IM_specific_down %>% mutate(id = trimws(as.character(id)))

# Annotation function, need to combine this to the one before
add_annotations <- function(df) {
  df %>%
    mutate(
      symbol = mapIds(org.At.tair.db, keys = id, column = "SYMBOL",
                      keytype = "TAIR", multiVals = "first"),
      entrez = mapIds(org.At.tair.db, keys = id, column = "ENTREZID",
                      keytype = "TAIR", multiVals = "first")
    )
}

#------------------------ H-Specific TRANSPORTERS ------------------------#
H_up_transporters <- H_specific_up %>%
  semi_join(transporter_genes, by = "id") %>%
  left_join(transporter_genes, by = "id") %>%
  add_annotations()

H_down_transporters <- H_specific_down %>%
  semi_join(transporter_genes, by = "id") %>%
  left_join(transporter_genes, by = "id")%>%
  add_annotations()

# IM specific
IM_up_transporters <- IM_specific_up %>%
  semi_join(transporter_genes, by = "id") %>%
  left_join(transporter_genes, by = "id")%>%
  add_annotations()

IM_down_transporters <- IM_specific_down %>%
  semi_join(transporter_genes, by = "id") %>%
  left_join(transporter_genes, by = "id")%>%
  add_annotations()

#------------------------ CLEAN DATA MATRIX ------------------------#
# H
clean_rank_H <- function(df, direction = "up") {
  df %>% select(id, symbol, logFC_x, padj_x, Family, Substrate,
           Transport_Type = `Transport Type`,
           TC_Num = `TC Num`) %>%
        # remove Inf and -Inf values
    filter(!is.infinite(logFC_x)) %>%
    arrange(if (direction == "up") dplyr::desc(logFC_x) else logFC_x) %>%
    mutate(
      rank = row_number(),# add rank by log fold change
      padj = padj_x,
      FC_label = round(logFC_x, 2)# rounded fold change for table
    )}
H_up_transporters <- clean_rank_H(H_up_transporters, "up")
H_down_transporters <- clean_rank_H(H_down_transporters, "down")

# IM  
clean_rank_IM <- function(df, direction = "up")  {
  df %>% select(id, symbol, logFC_y, padj_y, Family, Substrate,
           Transport_Type = `Transport Type`,
           TC_Num = `TC Num`) %>%
        # remove Inf and -Inf values
    filter(!is.infinite(logFC_y)) %>%
    arrange(if (direction == "up") dplyr::desc(logFC_y) else logFC_y) %>%    
    mutate(
      rank = row_number(),  # add rank by log fold change
      padj = padj_y,
      FC_label = round(logFC_y, 2) # rounded fold change for table
    )}
IM_up_transporters <- clean_rank_IM(IM_up_transporters, "up")
IM_down_transporters <- clean_rank_IM(IM_down_transporters, "down")

#------------------------ COMBINE TRANSPORTERS ------------------------#
transporter_all <- bind_rows(
  H_up_transporters   %>% mutate(group = "H_Up"),
  H_down_transporters %>% mutate(group = "H_Down"),
  IM_up_transporters  %>% mutate(group = "IM_Up"),
  IM_down_transporters%>% mutate(group = "IM_Down")
)

#------------------------ CLEAN COMBINED MATRIX ------------------------#
transporter_all_clean <- transporter_all %>%
  mutate(
    logFC = case_when(
      group %in% c("H_Up", "H_Down") ~ logFC_x,
      group %in% c("IM_Up", "IM_Down") ~ logFC_y
    ),
    direction = ifelse(grepl("Up", group), "Up", "Down"),
    condition = ifelse(grepl("^H", group), "H", "IM"),
    padj = padj
  ) %>%
  filter(!is.na(logFC))  # safety filter

#------------------------ SAVE ------------------------#
Transporter_workbook <- list(
  H_up_transporters   = H_up_transporters,
  H_down_transporters = H_down_transporters,
  IM_up_transporters  = IM_up_transporters,
  IM_down_transporters= IM_down_transporters,
  transporter_all     = transporter_all
)

#writexl::write_xlsx(Transporter_workbook, file.path(tables_dir, "transporters_IM_H.xlsx"))
```

```{r}

# ---- KEGG: H vs UM ----
ekegg_HUM_up <- enrichKEGG(
  gene          = H_up_transporters$id,
  organism      = "ath",
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05
)

ekegg_HUM_down <- enrichKEGG(
  gene          = H_down_transporters$id,
  organism      = "ath",
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05
)

# ---- KEGG: IM vs UM ----
ekegg_IMum_up <- enrichKEGG(
  gene          = IM_up_transporters$id,
  organism      = "ath",
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05
)

ekegg_IMum_down <- enrichKEGG(
  gene          = IM_down_transporters$id,
  organism      = "ath",
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.05
)

# ---- Load required packages ----
library(clusterProfiler)
library(enrichplot)

# ---- Combine all KEGG enrichment results ----
combined_KEGG_all <- merge_result(list(
  HUM_up  = ekegg_HUM_up,
  IMUM_up = ekegg_IMum_up,
  HUM_down   = ekegg_HUM_down,
  IMUM_down   = ekegg_IMum_down
))

# ---- Visualize combined KEGG enrichment ----
dotplot(combined_KEGG_all,
        showCategory = 15,
        title = "Combined KEGG Enrichment: H vs UM and IM vs UM") +
  theme(axis.text.y = element_text(size = 10))
```


```{r}
#transporter_filtered <- transporter_all_clean %>%
 # group_by(Family) %>%
  #filter(n() >= 3) %>%     # keep only families with ≥ 3 genes
  #ungroup()

transporter_filtered <- transporter_all_clean %>%
  mutate(
    direction = factor(direction, levels = c("Up", "Down"))
  )

top_up <- transporter_filtered %>%
  filter(direction == "Up") %>%
  arrange(desc(logFC)) %>%
  slice_head(n = 10)

top_down <- transporter_filtered %>%
  filter(direction == "Down") %>%
  arrange(logFC) %>%
  slice_head(n = 10)

top_genes <- bind_rows(top_up, top_down) %>%
  mutate(
    condition = toupper(condition),
    direction = tolower(direction),
    group = paste(condition, direction, sep = "_")
  )

#writexl::write_xlsx(top_genes, file.path(tables_dir, "top_transporter_genes_H_IM.xlsx"))

```

```{r}
familyplot <- ggplot(transporter_filtered,
       aes(x = fct_reorder2(Family, direction, abs(logFC)),
           y = logFC,
           fill = condition)) +
 geom_col(position = position_dodge(width = 0.85),
         width = 0.75,
           color = "black",
           linewidth = 0.2) +
  geom_point(aes( alpha = padj, group = condition),
             position = position_dodge(width = 0.85),
             color = "black") +
  
  facet_wrap(~ direction, ncol = 1, scales = "free_y") +
  coord_flip() +
  
  scale_fill_manual(values = c("H" = "#CC79A7", "IM" = "#7BC8A4")) +
  scale_alpha_continuous(range = c(0.4, 1)) +
  scale_x_discrete(labels = function(x) {
  x[x == "The Multidrug/Oligosaccharidyl-lipid/Polysaccharide (MOP) Flippase Superfamily"] <- 
    "The Multidrug/Oligosaccharidyl-lipid/Polysaccharide\n(MOP) Flippase Superfamily"
  
  x[x == "The Proton-dependent Oligopeptide Transporter (POT) Family"] <- 
    "The Proton-dependent Oligopeptide Transporter\n (POT) Family"
  x
}) +
  labs(
    title = "Transporter Family log2FC Comparison (H vs IM)",
    subtitle = "Points reflect padj significance",
    x = "Transporter Family (ordered by |log2FC|)",
    y = "log2 Fold Change",
    fill = "Condition",
    alpha = "Significance (padj)"
  ) +
  
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 10, color = "gray30"),
    strip.text = element_text(face = "bold", size = 12),
    
  #gridlines
      panel.grid.major.y = element_line(color = "gray85"),
    panel.grid.major.x = element_line(color = "gray90"),
  #y axis
    axis.text.y = element_text(size = 11, color = "black", lineheight = 0.9),
  
    axis.text.x = element_text(size = 9, color = "black"),
    legend.position = "right"
  )

familyplot
#ggsave(file.path(graphics_dir, "transporter_family_plot.png"),
#       familyplot,
#       width = 14,
#       height = 10,
#       dpi = 300)
```


```{r}
library(ggplot2)
library(forcats)

ggplot(transporter_filtered,
       aes(x = fct_reorder(Family, logFC),
           y = logFC,
           fill = padj)) +
  geom_col() +
  facet_grid(direction ~ condition, scales = "free_y") +
  scale_fill_gradient2(
    mid = "steelblue",
    high = "firebrick",
    midpoint = 0
  ) +
  scale_size_continuous(range = c(1, 6)) +
  coord_flip() +
  labs(
    title = "Differential Expression of Transporter Families",
    x = "Transporter Family",
    y = "log2 Fold Change",
    fill = "log2FC"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold"),
    axis.text.y = element_text(size = 8),
    legend.position = "right"
  )

```

```{r}
toptransporter <- ggplot(top_genes,
       aes(x = fct_reorder(id, logFC),
           y = logFC,
           fill = group)) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values = c(
    "H_up" = "#CC79A7",
    "H_down" = "#F564E3",
    "IM_up" = "#7BC8A4",
    "IM_down" = "#4DAF7C"
  )) +
  labs(
    title = "Top Transporter Genes by log2 Fold Change",
    x = "Gene ID",
    y = "log2 Fold Change",
    fill = "Group"
  ) +
  theme_minimal(base_size = 12) +
    theme(
    strip.text = element_text(face = "bold"),
    axis.text.y = element_text(size = 10),
    legend.position = "right"
  )
toptransporter

ggsave(file.path(graphics_dir, "transporter_top_barplot.png"),
       toptransporter,
       width = 8,
       height = 8,
       dpi = 300)


```


# 21. TRANSPORTER HEATMAP
```{r TRANSPORTER HEATMAP}
# gene list
transporter_ids <- transporter_all$id
#------------------------ EXTRACT MATRIX ------------------------#
mat_trans <- vsd_mat[rownames(vsd_mat) %in% transporter_ids, ]

# scale per gene
mat_trans_scaled <- t(scale(t(mat_trans)))

# enforce row order
transporter_all <- transporter_all %>%
  mutate(group = factor(group, levels = c("H_Up", "H_Down", "IM_Up", "IM_Down")))

mat_trans_scaled <- mat_trans_scaled[transporter_all$id, ]

#------------------------ METADATA ------------------------#
metadata <- as.data.frame(colData(dds))
metadata$Groups <- factor(metadata$Groups, levels = c("UM", "IM", "H"))

# reorder columns
mat_trans_scaled <- mat_trans_scaled[, rownames(metadata[order(metadata$Groups), ])]
metadata <- metadata[order(metadata$Groups), , drop = FALSE]

#------------------------ HEATMAP ------------------------#
# column annotation
col_ha <- HeatmapAnnotation(
  Groups = metadata$Groups,
  col = list(
    Groups = c("UM" = "#56B4E9", "IM" = "#7BC8A4", "H" = "#CC79A7")
    ))

# row annotation
row_ha <- rowAnnotation(
  Group = transporter_all$group,
  col = list(
    Group = c(
      "H_Up"   = "steelblue",
      "H_Down" = "#DFFF00",
      "IM_Up"  = "#A52A2A",
      "IM_Down"= "orange")
  ))



# color scale
col_fun <- colorRamp2(c(-2, 0, 2), c("blue", "white", "red"))

transporter_heatmap <- Heatmap(
  mat_trans_scaled,
  name = "Z-score",
  col = col_fun,
  top_annotation = col_ha,
  left_annotation = row_ha,
  cluster_rows = FALSE,
  cluster_columns = TRUE,
  show_row_names = FALSE,
  show_column_names = TRUE,
  row_split = transporter_all$group,
  column_split = metadata$Groups,
  column_title = "Sample Groups: UM, IM, H",
  row_title = "Transporter Gene Groups"
)

#pdf(file.path(graphics_dir, "Transporter_Heatmap.pdf"), width = 10, height = 10)
transporter_heatmap
```

```{r}
plot_ranked_barplot <- function(df, title = "Ranked Transporters") {

  ggplot(df, aes(x = rank, y = logFC_x, fill = Family)) +
    geom_col() +
    geom_text(aes(label = FC_label), 
              vjust = -0.3, size = 3) +
scale_fill_manual(values = hcl.colors(
  length(unique(df$Family)), "Dark 3"
))+
    labs(
      title = title,
      x = "Rank (by logFC)",
      y = "log2 Fold Change",
      fill = "Family"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
          legend.text  = element_text(size = 8),   # smaller legend labels
      legend.title = element_text(size = 10)   # legend title
   
    )
}
plot_ranked_barplot(H_up_transporters, 
                    title = "Haustoria Upregulated Transporters")
```


```{r}
ggplot(transporter_all,
       aes(x = ifelse(grepl("H", group), logFC_x, logFC_y),
           y = padj,
           color = group)) +
  geom_point(alpha = 0.7) +
  facet_wrap(~ Family) +
  theme_minimal()
```


```{r}
ggplot(transporter_all,
       aes(x = Family,
           y = ifelse(grepl("H", group), logFC_x, logFC_y),
           fill = group)) +
  geom_col(position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # rotate and change the group color to trt, rm fewer than 2 or more genes. most to least order, color tissue localization, re order. hi to lo abs val expression order.
```


```{r}
ggplot(transporter_all,
       aes(x = rank,
           y = ifelse(grepl("H", group), logFC_x, logFC_y),
           fill = Family)) +
  geom_col() +
  facet_wrap(~ group, scales = "free") +
  theme_minimal()

```




```{r}
family_summary2 <- H_down_transporters %>%
  distinct(id, Family, logFC_x, log10_padj) %>%   # avoid duplicates
  group_by(Family) %>%
  summarise(
    gene_count = n(),
    median_logFC = median(logFC_x),
    median_log10_padj = median(log10_padj),
    .groups = "drop"
  ) %>%
  arrange(desc(gene_count))
# filter out very small families (e.g., 1 gene)
family_summary2 <- family_summary2 %>% filter(gene_count >= 2)

fam_plot <- ggplot(family_summary2,
                   aes(x = fct_reorder(Family, gene_count),
                       y = gene_count)) +
  
  # Bars = number of genes per family
  geom_col(fill = "grey85") +
  
  # Overlay bubbles = expression + significance
  geom_point(aes(size = median_log10_padj,
                 color = median_logFC)) +
  
  # Flip axes for readability
  coord_flip() +
  
  # Color gradient: blue = down, white = neutral, red = up
  scale_color_gradient2(
    low = "purple",   # lower induction
    mid = "white",
    high = "hotpink",  # stronger induction
    midpoint = 0
  ) +
  
  # Labels
  labs(
    title = "Transporter Family-Level Response in Haustoriated Cells",
    x = "Transporter Family",
    y = "Number of Genes",
    color = "Median log2FC",
    size = "Median -log10(padj)"
  ) +
  
  # Theme adjustments
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text.y = element_text(size = 8)
  )

# Display plot
print(fam_plot)


# Save the plot


#ggsave(file.path(graphics_dir, "H_down_transporter_family_plot.png"),
#       fam_plot,
#       width = 10,
#       height = 6,
#       dpi = 300)


# save the family summary table

#write.csv(family_summary2,
#          file.path(tables_dir, "H_down_transporter_family_summary.csv"),
#          row.names = FALSE)

#write_xlsx(family_summary2, file.path(tables_dir, "H_down_transporter_family_summary.xlsx"))

```

```{r}
library(dplyr)

# Combine Up and Down transporter families
H_up_summary <- H_up_transporters %>%
  distinct(id, Family, logFC_x, log10_padj) %>%
  mutate(Direction = "Up") %>%
  group_by(Family, Direction) %>%
  summarise(
    gene_count = n(),
    median_logFC = median(logFC_x),
    median_log10_padj = median(log10_padj),
    .groups = "drop"
  )

H_down_summary <- H_down_transporters %>%
  distinct(id, Family, logFC_x, log10_padj) %>%
  mutate(Direction = "Down") %>%
  group_by(Family, Direction) %>%
  summarise(
    gene_count = n(),
    median_logFC = median(logFC_x),
    median_log10_padj = median(log10_padj),
    .groups = "drop"
  )

combined_summary <- bind_rows(H_up_summary, H_down_summary)
combined_summary <- combined_summary %>% filter(gene_count >= 2)

# Make Direction a factor so Up is on top
combined_summary$Direction <- factor(combined_summary$Direction, levels = c("Up", "Down"))


```

```{r}
library(ggplot2)
library(forcats)

ggplot(combined_summary,
       aes(x = gene_count, y = fct_reorder(Family, gene_count),
           fill = Family)) +
  
  geom_col() +  
 scale_fill_viridis_d(option = "plasma", guide="none") +
  facet_wrap(~Direction, ncol = 1, scales = "free_y") +   # Up on top, Down on bottom
  scale_x_continuous(breaks = seq(0, 10, 2), limits = c(0,10)) +  # set ticks from 0-10 by 2
  labs(
    title = "Transporter Family-Level Response in Haustoriated Cells",
    x = "Number of Genes",
    y = "Transporter Family",
    fill = "Family"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
     strip.text.y = element_text(face = "bold", size = 8),
      axis.text = element_text(color = "black"),
             axis.text.y = element_text(size = 8),
            axis.text.x = element_text(size = 8), 
      legend.position = "right",
      strip.placement = "outside"
    )


ggsave(file.path(graphics_dir, "H_combined_families.png"),
      width = 8,
       height = 6,
       dpi = 300)
```

```{r}
summarize_families <- function(df, direction){
  df %>%
    distinct(id, Family, logFC_x, log10_padj) %>%
    mutate(Direction = direction) %>%
    group_by(Family, Direction) %>%
    summarise(
      gene_count = n(),
      median_logFC = median(logFC_x),
      median_log10_padj = median(log10_padj),
      .groups = "drop"
    )
}

H_up_summary <- summarize_families(H_up_transporters, "Up")
H_down_summary <- summarize_families(H_down_transporters, "Down")
combined_summary <- bind_rows(H_up_summary, H_down_summary)
combined_summary$Direction <- factor(combined_summary$Direction, levels = c("Up", "Down"))

family_order <- combined_summary %>%
  group_by(Family) %>%
  summarise(total_genes = sum(gene_count)) %>%
  arrange(desc(total_genes)) %>%
  pull(Family)



combined_summary$Family <- factor(combined_summary$Family, levels = family_order)
ggplot(combined_summary,
       aes(x = gene_count, y = Family)) +
  geom_col(fill = "grey85") +
  geom_point(aes(size = median_log10_padj, color = median_logFC)) +
  scale_color_gradient2(low = "purple", mid = "white", high = "hotpink", midpoint = 0) +
  facet_wrap(~Direction, ncol = 1, scales = "free_y") +
    scale_x_continuous(breaks = seq(0, 10, 2), limits = c(0,10)) +  # set ticks from 0-10 by 2
  labs(
    title = "Transporter Family-Level Response in Haustoriated Cells",
    x = "Number of Genes",
    y = "Transporter Family",
    color = "Median log2FC",
    size = "Median -log10(padj)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text.y = element_text(size = 8)
  )
ggsave(file.path(graphics_dir, "H_combined_families.png"),
      width = 10,
       height = 6,
       dpi = 300)
```



